Effects of Alternate Wetting and Drying Irrigation and Nitrogen Fertilization on Sheath Blight of Rice

### 2017-04-07 reports/Effects_of_AWD_N_on_SB.rmd
Edit abstract for clarity; fix spelling in M&M

Water and nitrogen management play vital roles in rice production. However, the mismanagement of these two management practices may trigger plant disease epidemics such as sheath blight of rice, caused by _Rhizoctonia solani_, which is favored by wet conditions, high relative humidity, and high nitrogen fertilizer levels. To understand how different combinations of water and nitrogen management affect sheath blight epidemics, we conducted two separate split-plot experiments with two irrigation regimes and differing nitrogen treatments in the dry seasons of 2015 and 2016. Disease scoring was the same in both experiments using a sheath blight assessment scale for field evaluation developed at the International Rice Research Institute to assess the severity on infected sheaths and leaves while sheath blight incidence on tillers were counted per hill. While we were unable to detect any differences in disease in either experiment due to irrigation methods, N rates or the interaction of the two treatments in either season we suggest that further research on this subject is warranted.

@@

# Materials and Methods
## Experimental design
Two experiments were conducted at the International Rice Research Institute's (IRRI) Ziegler Experiment Station in Los Baños, Calabarzon, Philippines (latitude 14.14˚, longitude 121.27˚) in the 2015 and 2016 dry seasons. For the 2016 season changes were made to optimize the experiment based on findings from the 2015 season. Both seasons consisted of split plot design with four replicates where irrigation was the main plot and N rate was the split plot treatent. The changes are detailed following.

>>>>

Water and nitrogen management play vital roles in rice production. However, the mismanagement of these two management practices may trigger sheath blight of rice, caused by _Rhizoctonia solani_, which is favored by wet conditions, high relative humidity, and high nitrogen fertilizer levels. To understand how different combinations of water and nitrogen management affect sheath blight epidemics, we conducted two separate split-plot experiments with a water saving (alternate wetting and drying) regime and traditional flood irrigation regime combined with differing nitrogen treatments in the dry seasons of 2015 and 2016. Disease was scored in the same way in both experiments using a sheath blight assessment scale for field evaluation developed at the International Rice Research Institute to assess the severity on infected sheaths and leaves while sheath blight incidence on tillers were counted per hill. We were unable to detect any differences in disease in either experiment due to irrigation regime, N rates or the interaction of the two treatments in either season. This suggests that farmers can adopt water saving technologies without risking increased sheath blight incidence. We suggest that further cross-cutting research in this area is warranted.

@@

# Materials and Methods
## Experimental design
Two experiments were conducted at the International Rice Research Institute's (IRRI) Ziegler Experiment Station in Los Baños, Calabarzon, Philippines (latitude 14.14˚, longitude 121.27˚) in the 2015 and 2016 dry seasons. For the 2016 season changes were made to optimize the experiment based on findings from the 2015 season. Both seasons consisted of split plot design with four replicates where irrigation was the main plot and nitrogen (N) rate was the split plot treatment. The changes between seasons and experiments are detailed following.
